Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{ I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback. }

### 2023-06-26 paper/paper.tex
#29 tighten existing discussion

\title{Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback. I thank Matthias Weigand in particular for providing me with his own data and for pointing me towards the paper by \cite{schoenholzer2022}.
}
}
\author{Elias Hadj Ammar}
\date{June 27, 2023}

@@

The results in Table \ref{tab:SW22_replication_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run. However, the second part of my analysis yields a null result: Table \ref{tab:baseline_50y} shows that the the positive overall effect we saw in Table \ref{tab:SW22_replication_50y} was not in fact driven by conquests. I do not find evidence in support of the hypothesis that selection for state power is responsible for the positive effect of switching.

I test the robustness of this second result by estimating Equation \eqref{eq:baseline} with different control variables $\textbf{X}_{it}$. We can see in Table \ref{tab:controls_50y} shows estimates of  obtained with different sets of control variables. The outcome is $All \, construction$ in all regressions. Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period. In Column (2) I remove this control. Columns (3) and (4) add conflict as a control variable, once in addition to switching and once without.

Comparing Column (2) to (1), and (4) to (3), we can see that it makes no difference whether we control for switches in the same period or not; the estimates are barely distinguishable. When we control for conflict, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude. If we trust these regressions, the probability of observing construction in a given year increases by 15 percentage points \textit{less} after conquests relative to successions.

At first glance this may suggest that conflict is an omitted variable in Column (1), biasing the baseline estimate. A closer look at the coefficients on $Conflict$ in Columns (3) and (4) makes this seem unlikely, though: they are very close to zero and not significant by any stretch of the imagination. For conflict to be an omitted variable it would need to be correlated both with the post-$Conquest$ dummy and construction. The first correlation seems plausible --- once-conquered cities may continue to be highly contested, which would imply a positive relationship, or they may enjoy protection from further attacks by their powerful new owner, which would imply a negative relationship --- but the second correlation is zero in my sample.\footnote
{
The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations. Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of -0.01.
} The difference cannot be the result of omitted variable bias.

I investigate one alternative explanation: because there are many missing values in the $Conflict$ variable, more than two thirds of observations are dropped between Columns (2) and (3). If the distribution of missing values across observations is not random, the different estimates may be due to starkly different samples. Figure \ref{fig:conflict_map} visualises how including conflict in the regression affects the distribution of the sample across space. The colour of each region codes the share of observations from the baseline sample that remain after dropping observations with missing conflict data. If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade. Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.

>>>>

\title{Conflict Driven Evolution and Urban Development: \\ Evidence from the Holy Roman Empire\footnote{
I thank Lara Andermann, Leonie Bielefeld, Daniel Bursian, Davide Cantoni, Emilio Esguerra, Filip Milojevic, Johannes Vorbach and Matthias Weigand for helpful comments and feedback.
}
}
\author{Elias Hadj Ammar}
\date{June 27, 2023}

@@

The results in Table \ref{tab:SW22_replication_50y} corroborate part of my hypothesis: switching states within the Holy Roman Empire had a positive effect on the development of cities in the long run. However, the null result in the rest of the analysis fails to convince us that conquests drove the positive overall effect. I find weak evidence at best for the hypothesis that selection for state power played a role in the economic development of European cities. In this section I check whether the null result in Table \ref{tab:baseline_50y} is robust to alternative specifications and discuss its meaning.

As a first check I include different controls $\textbf{X}_{it}$ in Equation \eqref{eq:baseline}. Table \ref{tab:controls_50y} shows the results of this exercise. Column (1) is identical to Column (1) in Table \ref{tab:baseline_50y} and controls only for switching in the same period. In Column (2) I remove this control. Columns (3) and (4) add conflict incidents as a control variable, once in addition to switching and once without. The outcome is $All \, construction$ in all regressions.

Comparing Column (2) to (1), and (4) to (3), we can see that it makes no difference whether we control for switches in the same period or not; the estimates are barely distinguishable. When we control for conflict incidents, however, the coefficient $\theta_{conquest}$ suddenly turns negative, highly significant and large in magnitude. After a conquest the probability of observing construction in a given period now increases by 15 percentage points \textit{less} than after a succession.

While at first this result may seem like evidence for omitted variable bias in the baseline estimate --- the "correct" estimate being the negative one in Column (3) --- a closer look at the coefficients on $Conflict$ makes this seem unlikely: they are close to zero and not significant by any stretch of the imagination. To cause bias conflict incidents would need to be correlated both with the post-$Conquest$ dummy and with construction, but the correlation with construction is empirically zero.\footnote
{
The coefficients on $Conflict$ in Table \ref{tab:controls_50y} are conditional correlations. Simply computing the correlation coefficient between $Conflict$ and $All\,construction$ yields a value of $-0.01$.
} 
Rather, I suggest that the coefficients are different because they are estimated on different samples: more than two thirds of observations are lost between Columns (2) and (3) because of missing values in the $Conflict$ variable. Figure \ref{fig:conflict_map} shows that the remaining observations follow quite a different distribution across space: The colour of each region codes the share of observations from the baseline sample that are included in the conflict sample. If the spatial composition of the sample stayed exactly the same, the entire map would be a uniform shade. Instead we immediately see big discrepancies: Thuringia and most northern regions all but disappear from the sample, while the more darkly shaded west and south retain between 30\% and 60\% of their observations.

If the difference between the coefficients in (1) and (3) is due to the different spatial distribution of the sample, that has the interesting implication of spatially heterogeneous effects. BLANK.
